---
title: "Using database"
author: ""
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}

```


### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2021, and select county with less than 10 pct vaccination

```{r}

```
#### Ordering and aggregation

1. Order counties with the highest rate of vaccination on Aug 14, 2021
2. Calculate the state level vaccine count


```{r}

```


### Merging tables

- merge vaccine and demo table by fips
- (advanced) calculate per 1000 cases and deaths

```{r}

```

## dblyr

Now we will use `dbplyr`

### Get the object for each table

```{r}

```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}

```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 14 Aug, 2021, and select county with more than 10 percent vaccination rate

```{r}

```
#### Ordering and aggregation

1. Order counties with the highest number of vaccine percentage
2. Calculate the state level vaccination total


```{r}

```


### Merging tables

- merge vaccine and demo table by fips
- calculate vaccination rate manually from Series_Complete_Yes and TotalPop
  - compare with Series_Complete_Pop_Pct


```{r}

```
