---
title: "Using database"
author: ""
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
dbGetQuery(db, "select * from demo limit 5")
dbGetQuery(db, "select * from trump limit 5")
dbGetQuery(db, "select * from vac limit 5")

dbGetQuery(db, "SELECT * FROM sqlite_schema
WHERE type='table'
ORDER BY name;")
```


### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2021, and select county with less than 10 pct vaccination

```{r}
dbGetQuery(db, "select * from vac where Date = '2022-07-27' limit 5")
dbGetQuery(db, "select * from vac where Date = '2021-08-01' and Series_Complete_Pop_Pct > 10")

```
#### Ordering and aggregation

1. Order counties with the highest rate of vaccination on Aug 14, 2021
2. Calculate the state level vaccine count


```{r}
df_temp <- dbGetQuery(db, "select Date, FIPS, Recip_County,  Recip_State, Series_Complete_Pop_Pct from vac where Date = '2021-08-14' order by Series_Complete_Pop_Pct desc")


```


### Merging tables

- merge vaccine and demo table by fips
- (advanced) calculate per 1000 cases and deaths

```{r}

dbGetQuery(db, "select vac.*, demo.FIPS from vac 
            inner join demo on vac.FIPS = demo.FIPS where Date = '2021-08-14' ")
```

## dblyr

Now we will use `dbplyr`

### Get the object for each table

```{r}
db_demo <- tbl(db, "demo")
db_vac <- tbl(db, "vac")
db_trump <- tbl(db, "trump")
```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
db_demo %>% 
  head()
db_trump %>% 
  head()

```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 14 Aug, 2021, and select county with more than 10 percent vaccination rate

```{r}
db_vac %>% count(Date) %>% arrange(desc(Date)) %>% head()
db_vac %>% filter(Date == "2022-07-27") %>%
  show_query()

df_new <- db_vac %>% filter(Date == "2022-07-27") %>% collect() 


df_new2 <- db_vac %>% filter(Date == "2021-08-14" & Series_Complete_Pop_Pct > 10) 
```
#### Ordering and aggregation

1. Order counties with the highest number of vaccine percentage
2. Calculate the state level vaccination total


```{r}
db_vac %>% filter(Date == "2022-07-27") %>% arrange(desc(Series_Complete_Pop_Pct)) %>% head(20)
#db_vac %>% filter(Date == "2022-07-27") %>% arrange(desc(Series_Complete_Pop_Pct)) %>% show_query()
```


### Merging tables

- merge vaccine and demo table by fips
- calculate vaccination rate manually from Series_Complete_Yes and TotalPop
  - compare with Series_Complete_Pop_Pct


```{r}
db_vac %>% filter(Date == "2022-07-27") %>%
  left_join(db_demo, by = "FIPS") %>% head()

db_vac %>% filter(Date == "2021-07-27") %>%
  left_join(db_demo, by = "FIPS") %>%
  mutate(vac_pct = Series_Complete_Yes / TotalPop) %>%
  select(Series_Complete_Pop_Pct, vac_pct) %>%
  collect() %>%
  drop_na() %>%
  cor()


```
